2H qSIP Calculations

### 2022-01-11 qSIP_calculations.Rmd
join compound abundance with gencalc data

## Import GC-FID Data
```{r}
root <- file.path("data", "gc_fid_data_2020")
file_list = list.files(path = root)
file_list <- file_list %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE) %>% 
  str_subset("BLK", negate = TRUE) # remove procedural blanks
sample_names <- file_list %>% 
  str_replace("FID.....", "") %>% 
  str_replace("_100ul-nhex.xls", "") %>% 
  str_replace("_100ul.xls", "")

@@

### Unnest FID integration details

```{r}
FID_results <- FID_all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # Specify chain length
  mutate(
    chain = `Peak Name` %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  # Rename these columns so we can join the IRMS and FID data
  # Important to specify what data comes from FID!
  rename(
    "sample_id" = sample,
    "compound" = `Peak Name`,
    "rt_FID" = `Retention Time`,
    "area_FID" = `Area`,
    "height_FID" = `Height`,
    "rel_area_FID" = `Relative Area`,
    "rel_height_FID" = `Relative Height`,
    "amount_FID" = `Amount`
         )
```

@@

```{r}
peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros %>% 
  left_join(FID_results, by = c("sample_id", "compound"))
```

@@

# Add extr efficiency back into results data frame
FID_results_abundance <- 
  left_join(
    FID_results_abundance,
    FID_results_abundance_23,
    by = "file"
  ) %>% 
  mutate(
    # Calculate amount of analyte in soil sample
    compound_ug_soil_sample = compound_ug_ext / extr_efficiency,
    # Calculate amount of analyte per gram of soil (3g was used for each extraction)
    soil_mass_g = 3,
    compound_ug_per_g_soil = compound_ug_soil_sample / soil_mass_g
  )
```

## Select only samples that are included in analysis

```{r}
# Remove meadow samples and Ag-ion fractions
FID_results_abundance_reported <- FID_results_abundance_no_stds %>% 
  filter(
    str_detect(sample_id, "m0", negate = TRUE), # remove meadow samples
    str_detect(sample_id, "m3", negate = TRUE), # remove meadow 30cm samples
    str_detect(sample_id, "t0t", negate = TRUE) # remove amended tundra samples
  )

# Uncomment to view included samples:
# FID_results_abundance_reported %>% select(sample_id) %>% unique() %>% view()
```

@@

```{r save_workspace_image}
# This will take a minute. 
# Uncomment below to save the workspace image.
save.image(file = "cache/qSIP_calculations.RData")
```

>>>>

## Import GC-FID Data
```{r}
root <- file.path("data", "gc_fid_data_2021")
file_list = list.files(path = root)
file_list <- file_list %>% 
  str_subset("blank", negate = TRUE) %>% 
  str_subset("Blank", negate = TRUE) %>% 
  str_subset("GCMS-STD", negate = TRUE) %>% 
  str_subset("BLK", negate = TRUE) # remove procedural blanks
sample_names <- file_list %>% 
  str_replace("FID.....", "") %>% 
  str_replace("_100ul-nhex.xls", "") %>% 
  str_replace("_100ul.xls", "")

@@

### Unnest FID integration details and clean

```{r}
FID_results <- FID_all_data %>% 
  select(-injection_details) %>% 
  unnest(integration_results) %>% 
  filter(nchar(`Peak Name`) > 0) %>% 
  # Specify chain length
  mutate(
    chain = `Peak Name` %>% 
      str_remove("\\d+-OH") %>% 
      parse_number() %>% 
      abs() %>% 
      {paste0("C", .)} %>% 
      factor()
      ) %>% 
  # Rename these columns so we can join the IRMS and FID data
  # Important to specify what data comes from FID!
  rename(
    "sample_id" = sample,
    "compound" = `Peak Name`,
    "rt_FID" = `Retention Time`,
    "area_FID" = `Area`,
    "height_FID" = `Height`,
    "rel_area_FID" = `Relative Area`,
    "rel_height_FID" = `Relative Height`,
    "amount_FID" = `Amount`
         ) %>% 
  # Clean up notation discrepencies
  mutate(
    compound = 
      case_when(
        str_detect(compound, "18:2 trans 9, 12") ~ "18:2 9, 12",
        str_detect(compound, "18:2 trans 9,12") ~ "18:2 9, 12",
        str_detect(compound, "18:2 9,12") ~ "18:2 9, 12",
        str_detect(compound, "16:1 cis-9") ~ "16:1 cis 9",
        str_detect(compound, "16:1 trans-9") ~ "16:1 trans 9",
        str_detect(compound, "18:1 cis-9") ~ "18:1 cis 9",
        str_detect(compound, "18:1 trans-9") ~ "18:1 trans 9",
        str_detect(compound, "21:0") ~ "21:0 (STD)",
        str_detect(compound, "23:0") ~ "23:0 (STD)",
        TRUE ~ compound
      )
  )
```

# Add extr efficiency back into results data frame
FID_results_abundance <- 
  left_join(
    FID_results_abundance,
    FID_results_abundance_23,
    by = "file"
  ) %>% 
  mutate(
    # Calculate amount of analyte in soil sample
    compound_ug_soil_sample = compound_ug_ext / extr_efficiency,
    # Calculate amount of analyte per gram of soil (3g was used for each extraction)
    soil_mass_g = 10, # 10 g of each soil were used for this study
    compound_ug_per_g_soil = compound_ug_soil_sample / soil_mass_g
  )
```

@@

```{r}
# peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros %>% 
#   left_join(FID_results, by = c("sample_id", "compound"))

peaks_mapped_with_zeros_FID <- peaks_mapped_with_zeros %>%
  left_join(
    FID_results_abundance_no_stds, 
    by = c("sample_id", "compound")
  )
```

@@

@@

```{r save_workspace_image}
# This will take a minute. 
# Uncomment below to save the workspace image.
# save.image(file = "cache/qSIP_calculations.RData")
```
